### 2016-10-09 paper.tex
Added motivation

\documentclass[]{article}

\usepackage{hyperref}

%opening
\title{Automatic extraction of network signatures from malware code}
\author{Arjun Gurumurthy, Kausik Subramanian}

\begin{document}

\maketitle

\begin{abstract}
Hello
\end{abstract}

\section{Motivation}
Malware, short for malicious software, 
is one of the primary threats to modern-day networks, especially
with the mass proliferation of the Internet. These are used to 
disrupt computer operations, extract sensitive information, and 
gain access to private systems.  Infected hosts can participate 
in distributed denial-of-service attacks (DDoS), these class of 
attacks are very hard to detect and mitigate, leading to immense
loss of revenues. A new shift is towards creating malware for 
profit called botnets, in which a customer can buy a network of
infected hosts for a price. In such a setting, the malware is
specifically designed to evade detection, thus motivating the need
for better detection tools for malware.

The most prevalent form of malware detection is endhost-based detection
(commonly known as anti-virus software). Detected malware's code is hashed
and the code signatures are uploaded to a central repository. By maintaining
an up-to-date repository of malware signatures, the endhost software can
detect malware by comparing program signatures with the repository. However, 
endhost-based malware detection can fail due to various reasons~\cite{networksig}:
\begin{itemize}
	\item \textbf{Anti-virus software is not operational or up to date.} 
	Due to this, the endhost machine can be infected by malware without any
	noticeable effects, thus the user will be oblivious of the infection.
	\item \textbf{Packer programs.} To render programs stealthy, malware authors
	employ packer programs~\cite{packer}. Packers change the program content so that
	its signature differs but its functionality has not changed, thus avoiding
	detection by endhost signature-based software.
\end{itemize}

Instead, a intrusion-detection mechanism integrated 
directly into the service provider’s network offers a
much-needed additional layer of protection. While the malware
code is different, the protocol of the malware used to talk to its
control server (which issues orders to the malware) will not vary. 
Thus, by generating accurate \emph{network signature} of the malware (which
can informally defined as the set of messages sent and received by the malware to
initiate an action), we can use the intrusion detection system in the
network (for e.g., Bro~\cite{bro}) to search for these particular signatures,
and can be used to infer the presence of the malware in the network. However,
manually inspecting malware code is an intractable approach (signatures may 
need to be extracted in real-time). Thus, we propose \emph{automatic extraction 
of accurate network signatures from malware code using different static
analysis techniques.}

\section{Challenges}

\bibliographystyle{abbrv}
\bibliography{references}

\end{document}

### 2016-10-10 paper.tex
Started challenges

@@
 \begin{abstract}
-Hello
+
 \end{abstract}
@@
 \section{Challenges}
+We abstract the problem to the following: given a program point,
+can we automatically provide an ordered set of symbolic packets 
+that must be received (and sent) by the client such that the program 
+reaches the program point. The symbolic packets can be used as
+network signatures and used to detect malware (a set of packets
+satisfying the symbolic packets in order will infer that there 
+is a malware with high confidence). By constructing symbolic packet
+signatures, we can generalise the intrusion policy and reduce 
+true negatives.
 

### 2016-10-10 paper.tex
finished proposal

@@
 %opening
-\title{Automatic extraction of network signatures from malware code}
+\title{Automatic extraction of symbolic network signatures from malware code}
 \author{Arjun Gurumurthy, Kausik Subramanian}
@@
 \begin{abstract}
-
+We propose the automatic extraction of symbolic network signatures from malware code to aid and automate network-based malware detection systems.
 \end{abstract}
@@
 signatures, we can generalise the intrusion policy and reduce 
-true negatives.
+true negatives (thus not affecting legitimate traffic).
 
+Malware code can be thought of as a program reacting based on input received
+packets received (possibly multi-threaded), which
+could require changes to existing reachability analyses tools 
+like Codesurfer~\cite{codesurfer}. To extract symbolic network
+signatures, we need to define the abstract semantics of the program
+suited for network signatures, and define the \emph{combine} and 
+\emph{extend} operators used for interprocedural analysis.
+
+\section{Project Proposal}
+\textbf{Tasks}:
+\begin{itemize}
+	\item Identify and characterize features of common malware code - 1-2 weeks
+	\item Develop the abstract semantics and operators for extracting  
+	symbolic network signatures from malware code - 2-3 weeks
+	\item Modifying/Extending existing tools like Codesurfer for symbolic 
+	network signature extraction - 2-3 weeks
+\end{itemize}
+
+\noindent\textbf{Open source tools}: Static analysis tools, malware repositories
+
+\noindent\textbf{Evaluation}: 
+The primary metrics of evaluation are time taken for analysis, accuracy and 
+generality of the symbolic signatures. Performance analysis can be performed 
+with different malware programs with varying lines of code to understand
+the time complexity of the analysis. The quality of the symbolic network
+signatures is a tougher metric to measure. Ideally, the symbolic network 
+signature must only include packets which will make the program reach the
+specified point. If the signature is more general, legitimate packets may 
+be considered as malicious and blocked. On the other hand, the signature
+may miss actual malware traffic because the signature does not encompass
+all inputs leading to the point. We would try to address these trade-offs in
+our evaluation.
 \bibliographystyle{abbrv}

### 2016-10-30 paper.tex
Added progress milestone 1

@@
 \documentclass[]{article}
+\usepackage[xcolor]{changebar} 
+\usepackage{hyperref}
+\usepackage{cleveref}
+\usepackage{listings}
+\newcommand{\deleted}[1]{%
+	\cbcolor{red}
+	\begin{changebar}
+		#1
+	\end{changebar}%
+}%
 
-\usepackage{hyperref}
+\newcommand{\added}[1]{%
+	\cbcolor{green}
+	\begin{changebar}
+		#1
+	\end{changebar}%
+}%
+\usepackage{tikz}
+\usetikzlibrary{automata,positioning,arrows}
 
@@
 	\item Identify and characterize features of common malware code - 1-2 weeks
-	\item Develop the abstract semantics and operators for extracting  
-	symbolic network signatures from malware code - 2-3 weeks
+	\added{Refer to ---}
+	\item \deleted{Develop the abstract semantics and operators for extracting  
+	symbolic network signatures from malware code - 2-3 weeks} \\
+	\added{Automatically generate a symbolic automaton of the given program
+		based on packet and variable state - 2-3 weeks}
 	\item Modifying/Extending existing tools like Codesurfer for symbolic 
@@
 our evaluation.
+
+\appendix
+\section{Malware program model}
+The malware program listens to incoming packets, and the
+network signature is the set of incoming packets which triggers
+an attack by the program. We propose a reasonable abstraction
+of the code of malware programs.
+
+We model a malware program as a infinite reactive system 
+of the form:
+\begin{lstlisting}	
+while(True) {
+	packet = socket.receive();
+	...
+}
+\end{lstlisting}
+Each iteration of the loop, the program receives a
+packet, and based on current state and packet received,
+it performs certain actions. Using static analysis
+techniques, we can identify two set of variables, \emph{state}
+and \emph{packet} variables. State variables represent the 
+program state in terms of the protocol, which packet variables
+are used to represent the current packet being processed by
+the program.
+\begin{lstlisting}	
+while(True) {
+	packet = socket.receive();
+	...
+	if (a == 1 and packet[8] = "b") {
+		a = a + 1;
+	}
+	...
+}
+\end{lstlisting}	
+In the above code snippet, $a$ is a state variable
+and $packet$ is a packet variable. This also
+illustrates the packet processing 
+function of the program, which based on the current
+input packet and state variables, either transitions
+to a new program state or performs an action (attack or
+send packets). 
+\begin{lstlisting}	
+a = 0;
+while(True) {
+packet = socket.receive();
+	if (a == 0 and packet[0] = "a") {
+		a = a + 1;
+	}
+	if (a == 1 and packet[0] = "b") {
+		a = a + 1;
+	}
+	if (a == 2 and packet[0] = "c") {
+		attack()  [the attack point is provided as input to analysis]
+	}	
+}
+\end{lstlisting}	
+The above code snippet represents the complete model
+of the malware program, where based on input packets and
+current state, the program transitions to a new state. 
+The symbolic automaton for this program is 
+shown in \Cref{fig:progautomaton}. Network signatures
+for the malware are all paths from start state ($a=0$)
+to state of attack. In this case, the sequence of packets
+in order to initiate an attack is $a* \rightarrow b* \rightarrow c*$.
+\begin{figure}[h]
+	\centering
+	\begin{tikzpicture}[shorten >=0.5pt,node distance=3cm,on grid,auto] 
+	\node[state, initial] (v0)   {$a = 0$}; 
+	\node[state] (v1) [right=of v0] {$a = 1$}; 
+	\node[state] (v2) [right=of v1] {$a = 2$};
+	\node[state] (v3) [right=of v2] {$attack$}; 	
+	\path[->] 
+	(v0) edge node {$a*$} (v1)
+	(v1) edge node {$b*$} (v2)
+	(v2) edge node {$c*$} (v3);
+	\end{tikzpicture}
+	\caption{Symbolic automaton for malware program.}
+	\label{fig:progautomaton}
+\end{figure}
 \bibliographystyle{abbrv}

### 2016-10-31 paper.tex
add related work

@@
 	\item Identify and characterize features of common malware code - 1-2 weeks
-	\added{Refer to ---}
+	\added{Refer to sections A and B}
 	\item \deleted{Develop the abstract semantics and operators for extracting  
@@
 	if (a == 2 and packet[0] = "c") {
-		attack()  [the attack point is provided as input to analysis]
+		attack()  [the attack point is provided as 
+				input to analysis]
 	}	
@@
 \end{figure}
+
+\section{Related Work}
+An important assumption that we make is that of the simplification of the botnet client code to a set of \textbf {if (state, packet)} statements.
+Looking at existing research, significant work has gone into automatic protocol reverse-engineering of botnet C\&C protocols using dynamic binary analysis based on network traces.
+This generally follows a two step process: Extract the protocol grammar, that captures the structure of messages comprising the protocol.
+Knowing the protocol message format would enable the construction of the protocol state machine, which represents the sequence of messages as a specification of protocol states and transition between states (based on messages received).
+For example, Polyglot~\cite{polyglot} is a system that uses dynamic binary analysis to automatically extract the protocol message format, and Dispatcher~\cite{dispatcher} is a tool that makes use of techniques specified in Polyglot to extract the message format and field semantics for messages sent and received by the application.
+
+Taint propagation is used by the tool to map how the packets received manipulate the state.
+When the program calls the \textit{read()} function, Dispatcher uses the function's prototype to determine if any argument is tainted based on the memory locations it accesses.
+Via this approach, a spam botnet called MegaD is analyzed by Dispatcher.
+
+Thus, having knowledge of the network packets received would enable us to make inferences on how they could transition to an 'attack point'.
+Our work focuses on static analysis of the malware code, without having to execute it and/or capture network traces.
+
+In addition, work has been done on inferring state machines automatically for stateful network protocols - Prospex~\cite{prospex}.
+Prospex essentially performs clustering of similar network messages and tracks protocol state changes.
+By merging similar states, a minimal state machine is produced, and the paper produces a labeled state tree for the Agobot malware.
+This could be used as our initial malware program model for which we determine the sequence of packets to initiate an attack.
+
 \bibliographystyle{abbrv}

### 2016-11-21 paper.tex
Progress report

--- 2016-10-30 paper.tex
@@
 \begin{itemize}
-	\item Identify and characterize features of common malware code - 1-2 weeks
-	\added{Refer to ---}
-	\item \deleted{Develop the abstract semantics and operators for extracting  
-	symbolic network signatures from malware code - 2-3 weeks} \\
-	\added{Automatically generate a symbolic automaton of the given program
-		based on packet and variable state - 2-3 weeks}
-	\item Modifying/Extending existing tools like Codesurfer for symbolic 
-	network signature extraction - 2-3 weeks
+	\item Identify and characterize features of common malware code - 1-2 weeks.
+	Refer to \Cref{sec:malwaremodel}
+	\item \added{Extracting a network signature of length $n$ 
+		from the program \\
+	Refer to \Cref{sec:extraction}
+	\item  Generalizing the network signature extraction and further optimizations
+	and evaluation.}
+	\item \deleted{Modifying/Extending existing tools like Codesurfer for symbolic 
+		network signature extraction - 2-3 weeks}
 \end{itemize}
@@
 \appendix
-\section{Malware program model}
+\section{Malware program model} \label{sec:malwaremodel}
 The malware program listens to incoming packets, and the
@@
 packet = socket.receive();
-	if (a == 0 and packet[0] = "a") {
+	if (a == 0 and packet[0] == "a") {
 		a = a + 1;
 	}
-	if (a == 1 and packet[0] = "b") {
+	if (a == 1 and packet[0] == "b") {
 		a = a + 1;
 	}
-	if (a == 2 and packet[0] = "c") {
+	if (a == 2 and packet[0] == "c") {
 		attack()  [the attack point is provided as input to analysis]
@@
 \end{figure}
-\bibliographystyle{abbrv}
+
+\section{Signature Extraction} \label{sec:extraction}
+We now describe a procedure to generate a network signature
+of length $n$ (length of the signature is the number of packets
+received to trigger attack) for a boolean malware program. 
+We use the SMT solver Z3~\cite{z3}
+to find valid network signatures. In this section, we describe
+the constraints generated for the SMT solver from the program
+such that the model generated by the solver can be used to 
+extract signatures.
+
+\subsection{Preliminaries}
+For a malware program $M$, let $S$ denote
+ the vector of boolean state variables in the program,
+  and $P$ denote the vector of boolean packet variables
+  in the program\footnote{
+  	For example, a 60 bit packet is a vector 
+  	of 60 boolean variables.}. Let the
+  initial values of the state variables be $S_{0}$.
+
+As described in the earlier section, the malware
+packet processing code is a set of \texttt{if} 
+statements whose predicates are over the set of state
+and packet variables. Let the set of \texttt{if} 
+statements be denoted by $I$, and let us denote each \texttt{if} 
+statement by $(\gamma, \delta) \in I$, where $\gamma(S,P)$
+denotes the predicate on $S$ and $P$, and $\delta(S,P)$ denotes
+the action on $S$ if $\gamma(S,P)$ is satisfied. Therefore,
+the semantics of the \texttt{if} statement is as follows:
+\begin{equation}
+\gamma(S,P) \Rightarrow S' = \delta(S,P)
+\end{equation}
+where $S'$ denotes the new state after the \texttt{if} 
+block is executed.
+
+\subsection{Transition Model}
+We denote the state
+vector after receiving $i$ packets as $S_i$.
+When packet $i+1$ of the signature is received, 
+mlaware $M$, based on $S_i$ and $P_{i+1}$, will execute 
+a \texttt{if} $(\gamma_j, \delta_j)$ block\footnote{
+	We assume one \texttt{if} block is executed for every packet.}.
+Therefore, we can model the malware state transition 
+by the following constraints: 
+\begin{equation}
+	\forall i. 0 \leq i \leq n-2.~\forall (\gamma, \delta) \in I. ~\gamma(S_i, P_{i+1}) \Rightarrow S_{i+1} = \delta(S_i, P_{i+1})
+\end{equation}
+Let the attack point block be $(\gamma_M, \delta_M)$. 
+Therefore, a network signature of length $n$ means
+that $P_n$ triggers the attack. The state of the program
+is $S_{n-1}$. The following constraint holds:
+\begin{equation}
+	\gamma_M(S_{n-1},P_{n})
+\end{equation}
+The SMT solver will return a model for packets $P_1,
+\ldots P_n$ such that the program after these receiving
+these packets reaches the attack point. Therefore, 
+we run this procedure for varying lengths $n$ to find
+network signatures of different lengths.
+
+Also, this procedure only returns a particular signature
+of length $n$. We can generalize it by analyzing the 
+transitions of $M$ to generalize the signature. For
+example, if the $i^{th}$ packet only matches certain 
+predicates in the transition, we can generalize the packet on
+these bits (the rest of the bits are don't care). To 
+generate the different singatures of length $n$, 
+we can run the SMT solver multiple times to get different 
+models satisfying the constraints.
+
+\subsection{Example}
+Let us consider the following program receiving two-bit packets:
+\begin{lstlisting}	
+s1 = 0;
+s2 = 0;
+while(True) {
+	packet = socket.receive();
+	if (s1 == 0 and packet[0] == 0) {
+	s1 = 1;
+	}
+	if (s1 == 1 and s2 == 0 and packet[1] == 1) {
+	s2 = (not s2) and packet[1];
+	}
+	if (s1 == 1 and s2 == 1 and packet[0] == 1 and packet[1] == 1) {
+	attack()  
+	}	
+}
+\end{lstlisting}	
+We denote the state variables by $s_1, s_2$ and packet 
+variables by $p_1, p_2$.
+For the first \texttt{if} block, 
+$\gamma: \neg s_1 \wedge \neg p_1$, and $\delta: s_1'$ ($s_1$ 
+is set by the action). Similarily,
+for the second block: 
+$\gamma: s_1 \wedge \neg s_2 \wedge p_2$, and $\delta: 
+s_2' \Leftrightarrow \neg s_2 \wedge p_1$.
+For this program, initial state $S_0 = (0, 0)$.
+The attack point predicate $\gamma_M$ is as follows:
+\begin{equation}
+	s_1 \wedge s_2 \wedge p_1 \wedge p_1
+\end{equation}
+Therefore, if we add the constraints as described in the 
+earlier section, for $n=3$, we would obtain a
+valid network signature. 
 \bibliography{references}

### 2016-11-21 paper.tex

@@
 \begin{itemize}
-	\item Identify and characterize features of common malware code - 1-2 weeks.
-	Refer to \Cref{sec:malwaremodel}
-	\item \added{Extracting a network signature of length $n$ 
-		from the program \\
-	Refer to \Cref{sec:extraction}
-	\item  Generalizing the network signature extraction and further optimizations
-	and evaluation.}
-	\item \deleted{Modifying/Extending existing tools like Codesurfer for symbolic 
-		network signature extraction - 2-3 weeks}
+	\item Identify and characterize features of common malware code - 1-2 weeks
+	\added{Refer to sections A and B}
+	\item \deleted{Develop the abstract semantics and operators for extracting  
+	symbolic network signatures from malware code - 2-3 weeks} \\
+	\added{Automatically generate a symbolic automaton of the given program
+		based on packet and variable state - 2-3 weeks}
+	\item Modifying/Extending existing tools like Codesurfer for symbolic 
+	network signature extraction - 2-3 weeks
 \end{itemize}
 
-\noindent\textbf{Open source tools}: Static analysis tools, malware repositories
+\noindent\textbf{Open source tools}: SMT solvers, malware repositories
 
@@
 	if (a == 2 and packet[0] == "c") {
-		attack()  [the attack point is provided as input to analysis]
+		attack()  [the attack point is provided as 
+				input to analysis]
 	}	
@@
 \end{figure}
+
+\section{Related Work}
+An important assumption that we make is that of the simplification of the botnet client code to a set of \textbf {if (state, packet)} statements.
+Looking at existing research, significant work has gone into automatic protocol reverse-engineering of botnet C\&C protocols using dynamic binary analysis based on network traces.
+This generally follows a two step process: Extract the protocol grammar, that captures the structure of messages comprising the protocol.
+Knowing the protocol message format would enable the construction of the protocol state machine, which represents the sequence of messages as a specification of protocol states and transition between states (based on messages received).
+For example, Polyglot~\cite{polyglot} is a system that uses dynamic binary analysis to automatically extract the protocol message format, and Dispatcher~\cite{dispatcher} is a tool that makes use of techniques specified in Polyglot to extract the message format and field semantics for messages sent and received by the application.
+
+Taint propagation is used by the tool to map how the packets received manipulate the state.
+When the program calls the \textit{read()} function, Dispatcher uses the function's prototype to determine if any argument is tainted based on the memory locations it accesses.
+Via this approach, a spam botnet called MegaD is analyzed by Dispatcher.
+
+Thus, having knowledge of the network packets received would enable us to make inferences on how they could transition to an 'attack point'.
+Our work focuses on static analysis of the malware code, without having to execute it and/or capture network traces.
+
+In addition, work has been done on inferring state machines automatically for stateful network protocols - Prospex~\cite{prospex}.
+Prospex essentially performs clustering of similar network messages and tracks protocol state changes.
+By merging similar states, a minimal state machine is produced, and the paper produces a labeled state tree for the Agobot malware.
+This could be used as our initial malware program model for which we determine the sequence of packets to initiate an attack.
 
@@
 valid network signature.
+
+\bibliographystyle{abbrv}
 \bibliography{references}

### 2016-12-13 signet.py
Corrected version of implementation

@@ -1,4 +1,5 @@
 from z3 import *
+import random
 from collections import defaultdict
 
 class Signet(object) :
@@ -79,35 +80,50 @@ class Signet(object) :
 		return (outputVar == rexpr)
 
 	def extract(self) :
-		self.program()
-		for siglen in range(1,5):
+		self.generateProgram(5,5,5)
+		for siglen in range(1,10):
 			self.z3Solver.reset()
 			self.initializeSATVariables(siglen)
 
 			# Initial State
 			initExp = self.generatePredicate(self.init, 0)
 			self.z3Solver.add(initExp)
-			
+			# print initExp
 			# Transitions
 			for n in range(siglen - 1) :
+				predExpList = []
 				for [predicate, actions] in self.ifBlocks :
 					predExp = self.generatePredicate(predicate, n)
 
 					actionExpList = []
-					for action in actions :
-						actionExpList.append(self.generateAction(action, n))
+					for var in range(self.state) : 
+						actionExists = False
+						for action in actions :
+							if action[0] == var:
+								actionExpList.append(self.generateAction(action, n))
+								actionExists = True
+								break
+
+						if not actionExists : 
+							actionExpList.append(self.stateVariables[var][n+1] == self.stateVariables[var][n])
 
 					actionExp = And(*actionExpList)
+					# print "Action exp", actionExp
 					self.z3Solver.add(predExp == actionExp)
+					predExpList.append(predExp)
+
+				self.z3Solver.add(Or(*predExpList))
 
 			attackExp = self.generatePredicate(self.attackpredicate, siglen - 1)
 			self.z3Solver.add(attackExp)
+			# print attackExp
 
 			# Solve for network signature
 			modelsat = self.z3Solver.check()
 
 			if modelsat == z3.sat : 
 				print "Extracted a signature of length ", siglen
+				# print self.z3Solver.model()
 				self.signature(self.z3Solver.model(), siglen)
 			else : 
 				print "No signatures of length ", siglen
@@ -125,19 +141,68 @@ class Signet(object) :
 				else : 
 					packet = packet + "*"	
 			signature.append(packet)
-		
+		print signature	
 		self.signatures.append(signature)
 
 	def blockingSignature(self, signature, siglen):
+		pass
+
[…]

### 2016-12-15 paper.tex
Updated sig extr

--- 2016-11-21 paper.tex
@@
-\documentclass[]{article}
+\documentclass[twocolumn, 11pt]{paper}
 \usepackage[xcolor]{changebar} 
@@
 \usepackage{listings}
-\newcommand{\deleted}[1]{%
-	\cbcolor{red}
-	\begin{changebar}
-		#1
-	\end{changebar}%
-}%
 
-\newcommand{\added}[1]{%
-	\cbcolor{green}
-	\begin{changebar}
-		#1
-	\end{changebar}%
-}%
 \usepackage{tikz}
@@
 
-\section{Project Proposal}
-\textbf{Tasks}:
-\begin{itemize}
-	\item Identify and characterize features of common malware code - 1-2 weeks.
-	Refer to \Cref{sec:malwaremodel}
-	\item \added{Extracting a network signature of length $n$ 
-		from the program \\
-	Refer to \Cref{sec:extraction}
-	\item  Generalizing the network signature extraction and further optimizations
-	and evaluation.}
-	\item \deleted{Modifying/Extending existing tools like Codesurfer for symbolic 
-		network signature extraction - 2-3 weeks}
-\end{itemize}
+%\section{Project Proposal}
+%\textbf{Tasks}:
+%\begin{itemize}
+%	\item Identify and characterize features of common malware code - 1-2 weeks
+%	\added{Refer to sections A and B}
+%	\item \deleted{Develop the abstract semantics and operators for extracting  
+%	symbolic network signatures from malware code - 2-3 weeks} \\
+%	\added{Automatically generate a symbolic automaton of the given program
+%		based on packet and variable state - 2-3 weeks}
+%	\item Modifying/Extending existing tools like Codesurfer for symbolic 
+%	network signature extraction - 2-3 weeks
+%\end{itemize}
+%
+%\noindent\textbf{Open source tools}: SMT solvers, malware repositories
+%
+%\noindent\textbf{Evaluation}: 
+%The primary metrics of evaluation are time taken for analysis, accuracy and 
+%generality of the symbolic signatures. Performance analysis can be performed 
+%with different malware programs with varying lines of code to understand
+%the time complexity of the analysis. The quality of the symbolic network
+%signatures is a tougher metric to measure. Ideally, the symbolic network 
+%signature must only include packets which will make the program reach the
+%specified point. If the signature is more general, legitimate packets may 
+%be considered as malicious and blocked. On the other hand, the signature
+%may miss actual malware traffic because the signature does not encompass
+%all inputs leading to the point. We would try to address these trade-offs in
+%our evaluation.
 
-\noindent\textbf{Open source tools}: Static analysis tools, malware repositories
-
-\noindent\textbf{Evaluation}: 
-The primary metrics of evaluation are time taken for analysis, accuracy and 
-generality of the symbolic signatures. Performance analysis can be performed 
-with different malware programs with varying lines of code to understand
-the time complexity of the analysis. The quality of the symbolic network
-signatures is a tougher metric to measure. Ideally, the symbolic network 
-signature must only include packets which will make the program reach the
-specified point. If the signature is more general, legitimate packets may 
-be considered as malicious and blocked. On the other hand, the signature
-may miss actual malware traffic because the signature does not encompass
-all inputs leading to the point. We would try to address these trade-offs in
-our evaluation.
-
-\appendix
 \section{Malware program model} \label{sec:malwaremodel}
@@
 current state, the program transitions to a new state. 
-The symbolic automaton for this program is 
-shown in \Cref{fig:progautomaton}. Network signatures
-for the malware are all paths from start state ($a=0$)
-to state of attack. In this case, the sequence of packets
-in order to initiate an attack is $a* \rightarrow b* \rightarrow c*$.
-\begin{figure}[h]
-	\centering
-	\begin{tikzpicture}[shorten >=0.5pt,node distance=3cm,on grid,auto] 
-	\node[state, initial] (v0)   {$a = 0$}; 
-	\node[state] (v1) [right=of v0] {$a = 1$}; 
-	\node[state] (v2) [right=of v1] {$a = 2$};
-	\node[state] (v3) [right=of v2] {$attack$}; 	
-	\path[->] 
-	(v0) edge node {$a*$} (v1)
-	(v1) edge node {$b*$} (v2)
-	(v2) edge node {$c*$} (v3);
-	\end{tikzpicture}
-	\caption{Symbolic automaton for malware program.}
-	\label{fig:progautomaton}
-\end{figure}
+%
+%The symbolic automaton for this program is 
+%shown in \Cref{fig:progautomaton}. Network signatures
+%for the malware are all paths from start state ($a=0$)
+%to state of attack. In this case, the sequence of packets
+%in order to initiate an attack is $a* \rightarrow b* \rightarrow c*$.
+%\begin{figure}[h]
+%	\centering
+%	\begin{tikzpicture}[shorten >=0.5pt,node distance=3cm,on grid,auto] 
+%	\node[state, initial] (v0)   {$a = 0$}; 
+%	\node[state] (v1) [right=of v0] {$a = 1$}; 
+%	\node[state] (v2) [right=of v1] {$a = 2$};
+%	\node[state] (v3) [right=of v2] {$attack$}; 	
+%	\path[->] 
+%	(v0) edge node {$a*$} (v1)
+%	(v1) edge node {$b*$} (v2)
+%	(v2) edge node {$c*$} (v3);
+%	\end{tikzpicture}
+%	\caption{Symbolic automaton for malware program.}
+%	\label{fig:progautomaton}
+%\end{figure}
 
+\section{Related Work}
+An important assumption that we make is that of the simplification of the botnet client code to a set of \textbf{if (state, packet)} statements.
+Looking at existing research, significant work has gone into automatic protocol reverse-engineering of botnet C\&C protocols using dynamic binary analysis.
+This generally follows a two step process: Extract the protocol grammar that captures the structure of messages comprising the protocol.
+Knowing the protocol message format would enable the construction of the protocol state machine, which represents the sequence of messages as a specification of protocol states and transition between states (based on messages received).
+
+Polyglot~\cite{polyglot} is a system that automatically extracts the protocol message format using dynamic binary analysis.
+The intuition behind the design of the system is that the  way  that  an  implementation  of  the protocol processes the received application data  reveals a wealth of information about the protocol.
+
+Dispatcher~\cite{dispatcher} attempts to perform protocol reverse-engineering for active botnet infiltration.
+Essentially, it is a tool that makes use of Polyglot to extract the message format and in addition infers the field semantics for messages sent and received by the application.
+Thus, having knowledge of the network packets received would enable us to make inferences on how they could transition to an 'attack point'.
+
+Prospex~\cite{prospex} is a system for protocol specification extraction for stateful network protocols - essentially, it reverse engineers the protocol state machine by clustering similar network messages and tracks protocol state changes.
+By merging similar states, a minimal state machine is produced.
+
+All three systems use dynamic analysis techniques - mainly taint propagation.
+However, this would mean the code would need to execute in a sandboxed environment or be a part of a honeypot system.
+Static analysis techniques would be a better approach.
+
+Our work focuses on static analysis of the malware code, without having to execute it or capture network traces.
+
+\section{Live variable analysis} \label{sec:livevariableanalysis}
+We obtain network signatures using packet and state variables, the complexity increasing with respect to the number of state variables.
+To reduce the number of state variables, we perform live variable analysis for the program up until the attack point - that is, live variable analysis is done assuming the program terminates after the attack point.
+If the attack point is within one or more loops, then the assumed termination point would be immediately after all the loops are closed.
+For our use case, we can treat the entire code till the assumed termination point as a single block.
+Any variables that are not live within this block are ignored, and only the remaining live variables are used as state variables for the signature extraction step.
+Identifying live variables boils down to solving the dataflow equations:
+$$ in[n] = use[n] \quad \bigcup  \quad (out[n] - def[n]) $$
+$$ out[n] = \bigcup_{S \: \epsilon \: succ[n]}  in[s] $$
+
+where $in$ and $out$ represent the live variables entering and leaving the edge in the control flow graph.
+$def$ and $use$ represent the variables defined and referenced in an expression represented by a node in the CFG.
 \section{Signature Extraction} \label{sec:extraction}
@@
 \begin{equation}
-\gamma(S,P) \Rightarrow S' = \delta(S,P)
+\gamma(S,P) \Rightarrow S' == \delta(S,P)
 \end{equation}
@@
 
+Formally, given $M$, we need to find a sequence of packets $p_0$,
+$p_1, \ldots p_{n-1}$ such that $M$ starting from initial
+state reaches the attack point when it receives the packets in
+that particular order. This sequence is packets is defined as a 
+valid network signature of the malware program.
+
 \subsection{Transition Model}
 We denote the state
-vector after receiving $i$ packets as $S_i$.
-When packet $i+1$ of the signature is received, 
-mlaware $M$, based on $S_i$ and $P_{i+1}$, will execute 
-a \texttt{if} $(\gamma_j, \delta_j)$ block\footnote{
-	We assume one \texttt{if} block is executed for every packet.}.
-Therefore, we can model the malware state transition 
+vector before receiving the $i^{th}$ packet as $S_i$.
+When packet $i$ of the signature is received, 
+mlaware $M$, based on $S_i$ and $P_{i}$, will execute 
+a \texttt{if} $(\gamma_j, \delta_j)$ block.
+For a given action block comprising  
+assignment statements, the logical formula corresponding to
+the assignments is: 
+$\delta(S,P) = s_1' \Leftrightarrow \delta_1(S,P) \wedge s_2' \Leftrightarrow  \delta_2(S,P) \Leftrightarrow  \wedge ~s_{|S|}' = \delta_{|S|}(S,P)$, 
+where $\delta_i$ corresponds to the right hand side
+expression for the assignment statement $s_i = exp$.
+If the action has no assignment for a state 
+variable $s_i$, $\delta_i = s_i$ (the value must 
+remain unchanged, so we need to add this clause).
+
+We can model the malware state transition 
 by the following constraints: 
 \begin{equation}
-	\forall i. 0 \leq i \leq n-2.~\forall (\gamma, \delta) \in I. ~\gamma(S_i, P_{i+1}) \Rightarrow S_{i+1} = \delta(S_i, P_{i+1})
+\forall i. 0 \leq i \leq n-2.~\forall (\gamma, \delta) \in I. ~\gamma(S_i, P_{i}) \Rightarrow S_{i+1} \Leftrightarrow \delta(S_i, P_{i})
 \end{equation}
+
+To find minimal signatures, every packet received by the malware 
+must trigger a state change, which translated to the constraint
+that atleast one of the if predicates must be satisfied: 
+\begin{equation}
+\forall i. 0 \leq i \leq n-2. \bigvee_{(\gamma, \delta) \in I} \gamma(S_i, P_{i})
+\end{equation}
+
 Let the attack point block be $(\gamma_M, \delta_M)$. 
 Therefore, a network signature of length $n$ means
-that $P_n$ triggers the attack. The state of the program
-is $S_{n-1}$. The following constraint holds:
+that $P_{n-1}$ triggers the attack. The state of the program
+is $S_{n-1}$. The following constraint must hold:
 \begin{equation}
-	\gamma_M(S_{n-1},P_{n})
+\gamma_M(S_{n-1},P_{n-1})
 \end{equation}
@@
 generate the different singatures of length $n$, 
-we can run the SMT solver multiple times to get different 
-models satisfying the constraints.
+we can add blocking clauses for the signature we obtained 
+to find a signature different from the current one.
 
@@
 For the first \texttt{if} block, 
-$\gamma: \neg s_1 \wedge \neg p_1$, and $\delta: s_1'$ ($s_1$ 
+$\gamma: \neg s_1 \wedge \neg p_1$, and $\delta: s_1' \wedge s_2' \Leftrightarrow s_2$ ($s_1$ 
 is set by the action). Similarily,
@@
 $\gamma: s_1 \wedge \neg s_2 \wedge p_2$, and $\delta: 
-s_2' \Leftrightarrow \neg s_2 \wedge p_1$.
+s_1'\Leftrightarrow s_1 \wedge s_2' \Leftrightarrow \neg s_2 \wedge p_1$.
 For this program, initial state $S_0 = (0, 0)$.
@@
 valid network signature.
+
+\section{Evaluation}
+We present some preliminary evaluation of signature extraction using
+a synthetic malware example. We use the Z3~\cite{z3} SMT solver 
+for solving the constraints generated for extracting signatures. 
+We consider a malware program which processes 128 bit packets (i.e.,
+128 packet boolean variables). The malware program contains $n$ state
+variables which are initialized to false. For every packet satisfying 
+a certain condition, one of the state variables is set to true, and the 
+malware program launches an attack when all state variables are 
+set to true. Therefore, the signature length is $n$. We show the 
+time to solve constraints to extract the signature with varying 
+signature lengths in \Cref{fig:eval}.
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{plot.png}
+	\caption{Time taken to extract a signature using the Z3 solver
+		with varying signature lengths.}
+	\label{fig:eval}
+\end{figure}
+
+\section{Future Work}
+One of our simplifying assumptions is that we know the attack 
+point in the malware. However, with large codebases, this may be 
+difficult to identify manually. Therefore, an important extension is 
+Identifying attack points automatically. Some initial approaches can
+be identifying statements that make accesses to restricted files or 
+running restricted commands (e.g., sudo). To detect propagation
+in worms, 
+we can find code which causes the malware to send itself on the 
+network.
+
+Another simplifying assumption is our malware program model 
+as a set of if blocks in a while loop. We found real malware programs with
+large codebases with mutliple files, thus packet processing can be 
+across several procedures, requiring inter-procedural analyses.
+Also, it may be difficult to find malware code in the wild, only binaries,
+so extraction of signatures for a binary would be useful.
+
+Another feature of malware code can be loops whose number 
+of iterations depend on the bits of the packet. Thus, we cannot 
+unroll the loop trivially, and require to find invariants for the loop 
+predicated on packet characteristics.
+
+Finally, the signature extraction can be used to verify  
+implementations of receivers such that they adhere to 
+a protocol specification.
+
+\bibliographystyle{abbrv}
 \bibliography{references}

### 2016-12-15 paper.tex
Modifying related work and adding live variable analysis section

--- 2016-11-21 paper.tex
@@
 \end{figure}
+
+\section{Related Work}
+An important assumption that we make is that of the simplification of the botnet client code to a set of \textbf{if (state, packet)} statements.
+Looking at existing research, significant work has gone into automatic protocol reverse-engineering of botnet C\&C protocols using dynamic binary analysis.
+This generally follows a two step process: Extract the protocol grammar that captures the structure of messages comprising the protocol.
+Knowing the protocol message format would enable the construction of the protocol state machine, which represents the sequence of messages as a specification of protocol states and transition between states (based on messages received).
+
+Polyglot~\cite{polyglot} is a system that automatically extracts the protocol message format using dynamic binary analysis.
+The intuition behind the design of the system is that the  way  that  an  implementation  of  the protocol processes the received application data  reveals a wealth of information about the protocol.
+
+Dispatcher~\cite{dispatcher} attempts to perform protocol reverse-engineering for active botnet infiltration.
+Essentially, it is a tool that makes use of Polyglot to extract the message format and in addition infers the field semantics for messages sent and received by the application.
+Thus, having knowledge of the network packets received would enable us to make inferences on how they could transition to an 'attack point'.
+
+Prospex~\cite{prospex} is a system for protocol specification extraction for stateful network protocols - essentially, it reverse engineers the protocol state machine by clustering similar network messages and tracks protocol state changes.
+By merging similar states, a minimal state machine is produced.
+
+All three systems use dynamic analysis techniques - mainly taint propagation.
+However, this would mean the code would need to execute in a sandboxed environment or be a part of a honeypot system.
+Static analysis techniques would be a better approach.
+
+Our work focuses on static analysis of the malware code, without having to execute it or capture network traces.
+
+\section{Live variable analysis} \label{sec:livevariableanalysis}
+We obtain network signatures using packet and state variables, the complexity increasing with respect to the number of state variables.
+To reduce the number of state variables, we perform live variable analysis for the program up until the attack point - that is, live variable analysis is done assuming the program terminates after the attack point.
+If the attack point is within one or more loops, then the assumed termination point would be immediately after all the loops are closed.
+For our use case, we can treat the entire code till the assumed termination point as a single block.
+Any variables that are not live within this block are ignored, and only the remaining live variables are used as state variables for the signature extraction step.
+Identifying live variables boils down to solving the dataflow equations:
+$$ in[n] = use[n] \quad \bigcup  \quad (out[n] - def[n]) $$
+$$ out[n] = \bigcup_{S \: \epsilon \: succ[n]}  in[s] $$
+
+where $in$ and $out$ represent the live variables entering and leaving the edge in the control flow graph.
+$def$ and $use$ represent the variables defined and referenced in an expression represented by a node in the CFG.
 

### 2016-12-15 paper.tex
Reorganizing and final paper

--- 2016-12-15 paper.tex
@@
 \begin{abstract}
-We propose the automatic extraction of symbolic network signatures from malware code to aid and automate network-based malware detection systems.
+Malware is one of the primary threats to modern-day networks.
+While the most prevalent form of malware detection is endhost-based (e.g., matching code hashes), a detection mechanism integrated directly into the network offers a much-needed additional layer of protection.
+Towards this goal, we propose automatic generation of accurate network signatures of malware  using program analysis techniques and SMT solvers.
+Informally, a network signature is the set of messages received by the malware to initiate attacks.
+These network signatures can be used by intrusion detection systems to detect malware traffic in the network.
+We present some preliminary work towards this goal.
 \end{abstract}
@@
 \section{Motivation}
-Malware, short for malicious software, 
-is one of the primary threats to modern-day networks, especially
-with the mass proliferation of the Internet. These are used to 
+Malware is software designed to inflict illegitimate action and affect data, hosts or networks. These are used to 
 disrupt computer operations, extract sensitive information, and 
 gain access to private systems. Infected hosts can participate 
-in distributed denial-of-service attacks (DDoS), these class of 
+in distributed denial-of-service attacks (DDoS), and these class of 
 attacks are very hard to detect and mitigate, leading to immense
@@
 The most prevalent form of malware detection is endhost-based detection
-(commonly known as anti-virus software). Detected malware's code is hashed
-and the code signatures are uploaded to a central repository. By maintaining
-an up-to-date repository of malware signatures, the endhost software can
-detect malware by comparing program signatures with the repository. However, 
-endhost-based malware detection can fail due to various reasons~\cite{networksig}:
+(commonly known as anti-virus software).
+The detected malware's code is hashed and the code signatures are uploaded to a central repository.
+By maintaining an up-to-date repository of malware signatures, the endhost software can detect malware by comparing program signatures with the repository.
+However, endhost-based malware detection can fail due to various reasons~\cite{networksig}:
 \begin{itemize}
@@
 is a malware with high confidence). By constructing symbolic packet
-signatures, we can generalise the intrusion policy and reduce 
+signatures, we can generalize the intrusion policy and reduce 
 true negatives (thus not affecting legitimate traffic).
@@
 the program.
-\begin{lstlisting}	
-while(True) {
-	packet = socket.receive();
-	...
-	if (a == 1 and packet[8] = "b") {
-		a = a + 1;
-	}
-	...
-}
-\end{lstlisting}	
+
+\begin{figure}[h]
+	\includegraphics[width=\columnwidth]{code.PNG}
+	\label{fig:code1}
+\end{figure}
 In the above code snippet, $a$ is a state variable
@@
 send packets). 
-\begin{lstlisting}	
-a = 0;
-while(True) {
-packet = socket.receive();
-	if (a == 0 and packet[0] == "a") {
-		a = a + 1;
-	}
-	if (a == 1 and packet[0] == "b") {
-		a = a + 1;
-	}
-	if (a == 2 and packet[0] == "c") {
-		attack()  [the attack point is provided as 
-				input to analysis]
-	}	
-}
-\end{lstlisting}	
+\begin{figure}[h]
+	\includegraphics[width=\columnwidth]{code2.PNG}
+	\label{fig:code2}
+\end{figure}
 The above code snippet represents the complete model
@@
 Looking at existing research, significant work has gone into automatic protocol reverse-engineering of botnet C\&C protocols using dynamic binary analysis.
-This generally follows a two step process: Extract the protocol grammar that captures the structure of messages comprising the protocol.
+A general process followed is as follows: Extract the protocol grammar that captures the structure of messages comprising the protocol.
 Knowing the protocol message format would enable the construction of the protocol state machine, which represents the sequence of messages as a specification of protocol states and transition between states (based on messages received).
@@
 by the following constraints:
+
+$\forall i. 0 \leq i \leq n-2.~\forall (\gamma, \delta) \in I.~$
 \begin{equation}
-\forall i. 0 \leq i \leq n-2.~\forall (\gamma, \delta) \in I. ~\gamma(S_i, P_{i}) \Rightarrow S_{i+1} \Leftrightarrow \delta(S_i, P_{i})
+\gamma(S_i, P_{i}) \Rightarrow S_{i+1} \Leftrightarrow \delta(S_i, P_{i})
 \end{equation}
@@
 these bits (the rest of the bits are don't care). To 
-generate the different singatures of length $n$, 
+generate the different signatures of length $n$, 
 we can add blocking clauses for the signature we obtained 
 to find a signature different from the current one.
+\subsection{Example}
+Let us consider a program receiving two-bit packets.
 
-\subsection{Example}
-Let us consider the following program receiving two-bit packets:
-\begin{lstlisting}	
-s1 = 0;
-s2 = 0;
-while(True) {
-packet = socket.receive();
-if (s1 == 0 and packet[0] == 0) {
-s1 = 1;
-}
-if (s1 == 1 and s2 == 0 and packet[1] == 1) {
-s2 = (not s2) and packet[1];
-}
-if (s1 == 1 and s2 == 1 and packet[0] == 1 and packet[1] == 1) {
-attack()  
-}	
-}
-\end{lstlisting}	
 We denote the state variables by $s_1, s_2$ and packet 
@@
 $\gamma: \neg s_1 \wedge \neg p_1$, and $\delta: s_1' \wedge s_2' \Leftrightarrow s_2$ ($s_1$ 
-is set by the action). Similarily,
+is set by the action). Similarly,
 for the second block: 
@@
 s_1'\Leftrightarrow s_1 \wedge s_2' \Leftrightarrow \neg s_2 \wedge p_1$.
+\begin{figure}[ht]
+	\includegraphics[width=\columnwidth]{code7.PNG}
+	\label{fig:code7}
+\end{figure}
 For this program, initial state $S_0 = (0, 0)$.
@@
 as a set of if blocks in a while loop. We found real malware programs with
-large codebases with mutliple files, thus packet processing can be 
+large codebases with multiple files, thus packet processing can be 
 across several procedures, requiring inter-procedural analyses.
